---
title: "Week 2 suggested solutions to exercises"
author: "Martin Johnsson"
date: "19 januari 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
library(readr)
```



# 1

Read data:

```{r}
mens_long_jump_records <- read_csv("mens_long_jump_records.txt", comment = "#")
mens_long_jump_year_best <- read_csv("mens_long_jump_year_best.txt", comment = "#")
```

We combine them, and plot together:

```{r}
combined_jump <- rbind(transform(mens_long_jump_records, type = "record"),
                       transform(mens_long_jump_year_best, type = "year_best"))
qplot(x = year, y = meters, colour = type, data = combined_jump)
```

Alternatively, we can make a plot with multiple layers.

```{r}
ggplot() + geom_line(aes(x = year, y = meters), data = mens_long_jump_records) +
  geom_point(aes(x = year, y = meters), data = mens_long_jump_year_best)
```

# 2

Read data:

```{r}
womens_long_jump_year_best <- read_csv("womens_long_jump_year_best.txt", comment = "#")
year_best <- merge(womens_long_jump_year_best, mens_long_jump_year_best,
                   by.x = "year", by.y = "year")
colnames(year_best)[2:3] <- c("women", "men")
```

1986 is missing in the men's table.

(Full disclosure: there is nothing special about this year. I dropped it on purpose.)

```{r}
subset(year_best, year == 1986)
subset(mens_long_jump_year_best, year == 1986)
subset(womens_long_jump_year_best, year == 1986)
```
Therefore, the women's result gets dropped to. But there is a fix:

```{r}
year_best_fixed <- merge(womens_long_jump_year_best, mens_long_jump_year_best,
                         by.x = "year", by.y = "year", all.x = TRUE)
colnames(year_best_fixed)[2:3] <- c("women", "men")
subset(year_best_fixed, year == 1986)
```

(It will still be dropped when we calculate correlations, though.)

We make a scatterplot:

```{r}
qplot(x = women, y = men, data = year_best_fixed)
```

And calculate correlation:

```{r}
cor(year_best_fixed$women, year_best_fixed$men, use = "pairwise")
cor.test(year_best_fixed$women, year_best_fixed$men)
```


# 3

We melt the data frame:

```{r}
year_best_longform <- melt(year_best_fixed, id.vars = "year")
head(year_best_longform)
```

Dimensions line up:

```{r}
nrow(year_best_fixed)
nrow(year_best_longform)
nrow(year_best_longform)/2
```


# 4

```{r}
trigonometric_table <- function(x) {
  data.frame(x,
             sin_x = sin(x),
             cos_x = cos(x))
}
```

Try it out:

```{r}
x <- seq(from = 0, to = 2 * pi, by = 0.1)
trigonometric_table(x)
```

