---
title: "Week 3 suggested solutions"
author: "Martin Johnsson"
date: "10 februari 2017"
output: html_document
---

```{r}
library(plyr)
library(ggplot2)
```


## 1

```{r}
draws <- rnorm(1000, 2, 1)
length(which(draws < 0 | draws > 4))/1000
```

We can get the probability of a number smaller or greather than a certain value from the cumulative distribution function.

```{r}
pnorm(0, mean = 2, sd = 1)
pnorm(4, mean = 2, sd = 1, lower.tail = FALSE)
```

Because of symmetry, they are the same, so:

```{r}
2 * pnorm(0, mean = 2, sd = 1)
```


## 2

```{r}
many_draws <- replicate(10000, rnorm(100, 2, 1), simplify = FALSE)
many_means <- ldply(many_draws, mean)
quantile(many_means$V1, c(0.05, 0.95))
```

The standard error of the mean is one over the square root of the number of samples (i.e. 0.1).

An approximate 95% confidence interval is plus and minus two standard errors of the mean around:

```{r}
(sem <- 1 / sqrt(100))
c(2 - 2 * sem, 2 + 2 * sem)
```


## 3

```{r}
x <- sample(c("raindrops", "roses", "whiskers"), 100, replace = TRUE)
y <- 1 + 2* ifelse(x == "raindrops", 1, 0) + 3 * ifelse(x == "whiskers", 1, 0) + rnorm(100, 0, 2)
favourite_things <- data.frame(x, y)
```

Plot:

```{r}
qplot(x = x, y = y, data = favourite_things, geom = "boxplot")
```

Anova:

```{r}
model <- lm(y ~ x, data = favourite_things)
drop1(model, test = "F")
```


## 4

Simulation and analysis functions:

```{r}
simulate_things <- function(n) {
  x <- rep(c("raindrops", "roses", "whiskers"), n)
  y <- 1 + 0.5 * ifelse(x == "raindrops", 1, 0) + ifelse(x == "whiskers", 1, 0) + rnorm(n)
  data.frame(x, y)
}
analyze_things <- function(things) {
  lm(y ~ x, data = things)
}
get_p_value <- function(model) {
  drop <- drop1(model, test = "F")
  drop$"Pr(>F)"[2]
}
```

A test drive:

```{r}
one_rep <- analyze_things(simulate_things(20))
get_p_value(one_rep)
```

Replicated larger data:

```{r}
sim_things_20 <- replicate(1000, analyze_things(simulate_things(20)), simplify = FALSE)
sim_p_20 <- llply(sim_things_20, get_p_value)
sum(sim_p_20 < 0.05)/1000
```

Replicated smaller data:

```{r}
sim_things_10 <- replicate(1000, analyze_things(simulate_things(10)), simplify = FALSE)
sim_p_10 <- llply(sim_things_10, get_p_value)
sum(sim_p_10 < 0.05)/1000

sim_things_7 <- replicate(1000, analyze_things(simulate_things(7)), simplify = FALSE)
sim_p_7 <- llply(sim_things_7, get_p_value)
sum(sim_p_7 < 0.05)/1000
```


## 5

When we want the output in one data frame, ldply will aggregate it much more nicely than replicate.

```{r}
sim_data <- ldply(1:6, function(i) transform(simulate_things(10), replicate = i))
qplot(x = x, y = y, data = sim_data, geom = "boxplot") + facet_wrap(~ replicate)
```

A summary table using:

```{r}
ddply(sim_data, c("replicate", "x"), summarise, average = mean(y), stdev = sd(y))
```



## 6

```{r}
coin_toss <- function(nsim) {
  successes <- rbinom(nsim, size = 1, p = 0.5)
  ifelse(successes == 1, "t", "h")
}
```



## 7 

See if you can find a more elegant solution than my rather clunky llply call!

```{r}
x <- runif(100, 0, 1)
y <- 1 + x * 3 + unlist(llply(x, function(x) {
  if (x < 0.6)
    rnorm(1, 0, 1)
  else
    rnorm(1, 0, 2)
}))
qplot(x, y) + geom_smooth(method = lm)
```


## 8

A function to perform the above simulation:

```{r}
simulate_regression <- function(cutoff) {
  x <- runif(100, 0, 1)
  y <- 1 + x * 3 + unlist(llply(x, function(x) {
  if (x < cutoff)
    rnorm(1, 0, 1)
  else
    rnorm(1, 0, 2)
  }))
  data.frame(x, y)  
}
```

Generate random cutoffs and apply function.

Adding names makes ldply automatically label each simulation with the .id column.

```{r}
cutoffs <- runif(10, 0, 1)
names(cutoffs) <- 1:10
simulated_regressions <- ldply(cutoffs, simulate_regression)
```

Plots:

```{r}
qplot(x = x, y = y, data = simulated_regressions) + facet_wrap(~ .id)
```

