---
title: "Week 3 suggested solutions"
author: "Martin Johnsson"
date: "10 februari 2017"
output: html_document
---

## 1

```{r}
draws <- rnorm(1000, 2, 1)
length(which(a < 0 | a > 4))/1000
```

We can get the probability of a number smaller or greather than a certain value from the cumulative distribution function.

```{r}
pnorm(0, mean = 2, sd = 1)
pnorm(4, mean = 2, sd = 1, lower.tail = FALSE)
```

Because of symmetry, they are the same, so:

```{r}
2 * pnorm(0, mean = 2, sd = 1)
```


## 2

```{r}
many_draws <- replicate(10000, rnorm(100, 2, 1), simplify = FALSE)
many_means <- ldply(many_draws, mean)
quantile(many_means$V1, c(0.05, 0.95))
```

The standard error of the mean is one over the square root of the number of samples (i.e. 0.1).

An approximate 95% confidence interval is plus and minus two standard errors of the mean around:

```{r}
(sem <- 1 / sqrt(100))
c(2 - 2 * sem, 2 + 2 * sem)
```


## 3

```{r}
x <- sample(c("raindrops", "roses", "whiskers"), 100, replace = TRUE)
y <- 1 + 2* ifelse(x == "raindrops", 1, 0) + 3 * ifelse(x == "whiskers", 1, 0) + rnorm(100, 0, 2)
favourite_things <- data.frame(x, y)
```

Plot:

```{r}
qplot(x = x, y = y, data = favourite_things, geom = "boxplot")
```

Anova:

```{r}
model <- lm(y ~ x, data = favourite_things)
drop1(model, test = "F")
```


## 4

Simulation and analysis functions:

```{r}
simulate_things <- function(n) {
  x <- sample(c("raindrops", "roses", "whiskers"), n, replace = TRUE)
  y <- 1 + 0.5 * ifelse(x == "raindrops", 1, 0) + ifelse(x == "whiskers", 1, 0) + rnorm(n)
  data.frame(x, y)
}
analyze_things <- function(things) {
  lm(y ~ x, data = things)
}
get_p_value <- function(model) {
  drop <- drop1(model, test = "F")
  drop$"Pr(>F)"[2]
}
```

A test drive:

```{r}
one_rep <- analyze_things(simulate_things(100))
get_p_value(one_rep)
```

Replicated larger data:

```{r}
sim_things_100 <- replicate(1000, analyze_things(simulate_things(100)), simplify = FALSE)
sim_p_100 <- llply(sim_things_100, get_p_value)
sum(sim_p_100 < 0.05)/1000
```

Replicated smaller data:

```{r}
sim_things_50 <- replicate(1000, analyze_things(simulate_things(50)), simplify = FALSE)
sim_p_50 <- llply(sim_things_50, get_p_value)
sum(sim_p_50 < 0.05)/1000

sim_things_10 <- replicate(1000, analyze_things(simulate_things(10)), simplify = FALSE)
sim_p_10 <- llply(sim_things_10, get_p_value)
sum(sim_p_10 < 0.05)/1000
```


## 5

```{r}
get_estimates_p <- function(model) {
  coefs <- coef(model)
  data.frame(raindrops = coefs[1],
             roses = coefs[1] + coefs[2],
             whiskers = coefs[1] + coefs[3],
             p = get_p_value(model))
}
```

Apply it:

```{r}
estimates_100 <- ldply(sim_things_100, get_estimates_p)
mean(estimates_100$raindrops)
mean(estimates_100$roses)
mean(estimates_100$whiskers)
sign_estimates_100 <- subset(estimates_100, p < 0.05)
mean(sign_estimates_100$raindrops)
mean(sign_estimates_100$roses)
mean(sign_estimates_100$whiskers)
```
That looked good. How about a small sample size?

```{r}
estimates_10 <- ldply(sim_things_10, get_estimates_p)
mean(estimates_10$raindrops)
mean(estimates_10$roses)
mean(estimates_10$whiskers)
sign_estimates_10 <- subset(estimates_10, p < 0.05)
mean(sign_estimates_10$raindrops)
mean(sign_estimates_10$roses)
mean(sign_estimates_10$whiskers)
```


